Energy and the CES production function

### 2015-08-14 Articles/Article_EcolEcon_2015/EcolEcon_2015.Rnw
Added helpful comment.
Now includes reference model in fitted graphs.

Unadjusted and quality-adjusted data for both Portugal and the United Kingdom 
are shown in Figure~\ref{fig:Historical Data Graph}.

@@

<<Fitted_and_Residuals>>=
filenameResampleCoeffs <- ""
fittedAndResidData <- function(model, id){
  # ******************************************
  # getData doesn't work for lm models that are created for CSP="Adhere to CSP".
  # After we implement specified values for alpha_1, alpha_2, and alpha_3 in our fitting functions, 
  # getData(model) will probably work.
  # At that time, change model$data to getData(model)
  df <- data.frame(select(model$data, Year)) 
  # By including [c(1:length(df$Year))] on the end, we 
  # add NAs (as appropriate) to the end of these columns
  # if data do not exist.
  df$fitted <- yhat(model)[c(1:length(df$Year))]
  df$Residuals <- resid(model)[c(1:length(df$Year))]
  # Now add the response, using the original variable name.
  df[[summands(model$formula, right = FALSE)]] <- response(model)[c(1:length(df$Year))]
  # Now add meta information
  id_list <- strsplit(id, split = "[.]")[[1]]
  len <- length(id_list)
  df$Country <- id_list[[1]]
  # Exponential models have only 1 item in the id list, namely country. Accommodate that length with if statements.
  df$flavor <- if (len > 1) (if (id_list[[2]] ==  "unadjusted") "Unadjusted" else "Quality-adjusted") else NA
  df$energy <- if (len > 2) (if (id_list[[3]] == "noE") "Without energy" else "With energy") else "Without energy"
  df$model <- if (len > 3) id_list[[4]] else "exp"
  df$nest <- if (len > 4) id_list[[5]] else NA
  return(df)  
}

temp_data_exp <- leaf_apply(modelsExp, class=c("lm"), f = function(model, id){
  # ****************************************************
  # Not all of the results from this code are not correct at this time.
  # The residuals are OK.
  # But, the yhat() call in fittedAndResidData() will not return the correct information from the CSP models.
  # This is because modelsCSP was not created with our typical functions.
  # To correct this problem, we need to add arguments to cdModel and cesModel such that 
  # the models produced by those functions will be in the correct form for yhat() function.
  return(fittedAndResidData(model, id))
}) %>% do.call(rbind, .)
temp_data_exp$CSP <- NA

temp_data_CSP <- leaf_apply(modelsCSP, class=c("plm", "lm"), f = function(model, id){
  # ****************************************************
  # Not all of the results from this code are not correct at this time.
  # The residuals are OK.
  # But, the yhat() call in fittedAndResidData() will not return the correct information from the CSP models.
  # This is because modelsCSP was not created with our typical functions.
  # To correct this problem, we need to add arguments to cdModel and cesModel such that 
  # the models produced by those functions will be in the correct form for yhat() function.
  return(fittedAndResidData(model, id))
}) %>% do.call(rbind, .)
temp_data_CSP$CSP <- "Adhere to CSP"

temp_data_noCSP <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(model, id){
  return(fittedAndResidData(model, id))  
}) %>% do.call(rbind, .)
temp_data_noCSP$CSP <- "Reject CSP"

temp_data <- plyr::rbind.fill( temp_data_exp, temp_data_CSP, temp_data_noCSP )

# Get the unadjusted, without-energy residuals from models that adhere to the CSP
# and store them in a data frame
# r_unadj_noE_CSP <- filter(temp_data, flavor=="Unadjusted" & energy=="Without energy" & CSP=="Adhere to CSP") %>% 
#   select(Year, Country, Residuals, model)
# colnames(r_unadj_noE_CSP)[colnames(r_unadj_noE_CSP)=="Residuals"] <- "r_unadj_noE"
# Merge to get the unadjusted, without-energy values next to each country, year, and model
# fitted_and_resid_data <- merge(temp_data, r_unadj_noE_CSP)
# Create a column containing ∆r, the difference between a residual value
# and the residual value in the same year for the unadjusted, no-energy case
# fitted_and_resid_data[["Dr"]] <- abs(fitted_and_resid_data[["Residuals"]]) - abs(fitted_and_resid_data[["r_unadj_noE"]])

# Get the reference residuals (from the exponential-only models) and store them in a data frame
r_ref <- filter(temp_data, model=="exp") %>% select(Year, Country, Residuals)
colnames(r_ref)[colnames(r_ref)=="Residuals"] <- "r_ref"
# Merge to get the reference values in each appropriate row
fitted_and_resid_data <- merge(temp_data, r_ref)
# Create a column containing ∆r, the difference between a residual value
# and the residual value in the same year for the reference case
fitted_and_resid_data[["Dr"]] <- abs(fitted_and_resid_data[["Residuals"]]) - abs(fitted_and_resid_data[["r_ref"]])

@@

<<Fitted and Residual Graph Functions>>=
fittedGraphCSP <- function(data){
  ggplot() +
    geom_line(data=filter(data, CSP=="Adhere to CSP"), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=CSP)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) + 
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}

fittedGraphNoCSP <- function(data){
  ggplot() +
    geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=iGDP)) +
    geom_line(data=data, mapping=aes(x=Year, y=fitted, linetype=energy)) +
    scale_linetype_manual(values=c(2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}
@

Both historical GDP and fitted GDP are shown in 
Figures~\ref{fig:Graphs_CESkle-fitted}--\ref{fig:Graphs_CESekl-fitted}.
%
<<Graphs_CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraphNoCSP(filter(fitted_and_resid_data, model=="CES" & CSP=="Reject CSP" & (nest=="kle" | nest=="kl")))
@
%
<<Graphs_CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraphNoCSP(filter(fitted_and_resid_data, model=="CES" & CSP=="Reject CSP" & (nest=="lek" | nest=="kl")))
@
%
<<Graphs_CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraphNoCSP(filter(fitted_and_resid_data, model=="CES" & CSP=="Reject CSP" & (nest=="ekl" | nest=="kl")))
@

>>>>

Unadjusted and quality-adjusted data for both Portugal and the United Kingdom 
are shown in Figure~\ref{fig:Historical_Data_Graph}.

@@

<<Fitted_and_Residuals>>=
filenameResampleCoeffs <- ""
fittedAndResidData <- function(model, id){
  # ******************************************
  # getData doesn't work for lm models that are created for CSP="Adhere to CSP".
  # After we implement specified values for alpha_1, alpha_2, and alpha_3 in our fitting functions, 
  # getData(model) will probably work.
  # At that time, change model$data to getData(model)
}

temp_data_exp <- leaf_apply(modelsExp, class=c("lm"), f = function(model, id){
  df <- data.frame(select(model$data, Year, Country))
  df$fittedExp <- exp(fitted(model)) # exp accounts for the fact that the model was fitted in log space.
  df$ResidualsExp <- resid(model)
  # df$model <- "exp"
  return(df)
}) %>% do.call(rbind, .)

temp_data_CSP <- leaf_apply(modelsCSP, class=c("plm", "lm"), f = function(model, id){
  df <- data.frame(select(model$data, Year, Country, iGDP))
  df$fitted <- exp(fitted(model)) # fitted is in log space. Return to regular space.
  df$Residuals <- resid(model) # residuals are in log space, and that is what we want.
  # Now add meta information
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$flavor <- if (id_list[[2]] ==  "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  return(df)
}) %>% do.call(rbind, .)
temp_data_CSP$CSP <- "Adhere to CSP"

temp_data_noCSP <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(model, id){
  df <- data.frame(select(model$data, Year)) 
  # By including [c(1:length(df$Year))] on the end, we 
  # add NAs (as appropriate) to the end of these columns
  # if data do not exist.
  df$fitted <- yhat(model)[c(1:length(df$Year))]
  df$Residuals <- resid(model)[c(1:length(df$Year))]
  # Now add the response, using the original variable name.
  df[[summands(model$formula, right = FALSE)]] <- response(model)[c(1:length(df$Year))]
  # Now add meta information
  id_list <- strsplit(id, split = "[.]")[[1]]
  len <- length(id_list)
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] ==  "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  return(df)  
}) %>% do.call(rbind, .)
temp_data_noCSP$CSP <- "Reject CSP"

temp_data_noExp <- plyr::rbind.fill( temp_data_CSP, temp_data_noCSP )

# Now merge the exponential-only model with the CSP and noCSP models. 
# Doing this, because the exponential-only model is kind-of like iY:
# it applies to all situations.
# temp_data <- merge(temp_data_noExp, temp_data_exp)
fitted_and_resid_data <- merge(temp_data_noExp, temp_data_exp)

# No longer using this code, because we have a different "reference" model.
# Get the unadjusted, without-energy residuals from models that adhere to the CSP
# and store them in a data frame
# r_unadj_noE_CSP <- filter(temp_data, flavor=="Unadjusted" & energy=="Without energy" & CSP=="Adhere to CSP") %>% 
#   select(Year, Country, Residuals, model)
# colnames(r_unadj_noE_CSP)[colnames(r_unadj_noE_CSP)=="Residuals"] <- "r_unadj_noE"
# Merge to get the unadjusted, without-energy values next to each country, year, and model
# fitted_and_resid_data <- merge(temp_data, r_unadj_noE_CSP)
# Create a column containing ∆r, the difference between a residual value
# and the residual value in the same year for the unadjusted, no-energy case
# fitted_and_resid_data[["Dr"]] <- abs(fitted_and_resid_data[["Residuals"]]) - abs(fitted_and_resid_data[["r_unadj_noE"]])

# # Get the reference residuals (from the exponential-only models) and store them in a data frame
# r_ref <- filter(temp_data, model=="exp") %>% select(Year, Country, Residuals)
# colnames(r_ref)[colnames(r_ref)=="Residuals"] <- "r_ref"
# # Merge to get the reference values in each appropriate row
# fitted_and_resid_data <- merge(temp_data, r_ref)
# # Create a column containing ∆r, the difference between a residual value
# # and the residual value in the same year for the reference case
# fitted_and_resid_data[["Dr"]] <- abs(fitted_and_resid_data[["Residuals"]]) - abs(fitted_and_resid_data[["r_ref"]])

# Get the reference residuals (from the exponential-only models) and store them in a data frame
# r_ref <- select(temp_data_exp, Year, Country, ResidualsExp)
# colnames(r_ref)[colnames(r_ref)=="Residuals"] <- "r_ref"
# Merge to get the reference values in each appropriate row
# fitted_and_resid_data <- merge(temp_data, r_ref)
# Create a column containing ∆r, the difference between a residual value
# and the residual value in the same year for the reference case
fitted_and_resid_data[["Dr"]] <- abs(fitted_and_resid_data[["Residuals"]]) - abs(fitted_and_resid_data[["ResidualsExp"]])

@@

<<Fitted and Residual Graph Functions>>=
fittedGraphNoCSP <- function(data){
  fittedGraphData <- select(data, Year, Country, iGDP, fitted, fittedExp, flavor, energy, CSP, nest)
  fittedGraphData <- melt(fittedGraphData, 
                     measure.vars = c("iGDP", "fitted", "fittedExp"), 
                     value.name = "value", 
                     variable.name = "variable")
  fittedGraphData$label <- with(fittedGraphData, paste(variable, energy))
  # Now collapse some of the information. Don't need multiple versions of the exponential model, e.g.
  fittedGraphData <- filter(fittedGraphData, label=="iGDP With energy" | label=="fittedExp With energy" |
                       label == "fitted With energy" | label == "fitted Without energy")
  fittedGraphData$label <- plyr::mapvalues(fittedGraphData$label, 
                                      from = c("iGDP With energy", 
                                               "fittedExp With energy",
                                               "fitted With energy",
                                               "fitted Without energy"), 
                                      to = c("GDP", 
                                             "Reference model",
                                             "With energy",
                                             "Without energy"))
  # Relevel factors
  fittedGraphData$label <- relevelFactor(as.factor(fittedGraphData$label), 
                                         c("GDP", "Reference model", "Without energy", "With energy"))
  # Make (and return) the graph
  ggplot() +
    geom_line(data=fittedGraphData, mapping=aes(x=Year, y=value, group=label, linetype=label)) +
    scale_linetype_manual(values=c(1,4,2,3), name="") +
    facet_grid(Country ~ flavor) +
    scale_x_continuous(breaks=c(1960, 1980, 2000)) +
    scale_y_continuous(breaks=c(1, 2, 4, 6, 8, 10, 12)) +
    xlab("") +
    ylab("Indexed GDP (1960=1)") +
    xy_theme()
}
@

Both historical GDP and fitted GDP are shown in 
Figures~\ref{fig:Graphs_CESkle-fitted}--\ref{fig:Graphs_CESekl-fitted}.
%
<<Graphs_CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with $kl$ and ($kl$)$e$ nesting that reject the CSP. Solid line is historical GDP.">>=
fittedGraphData_kle <- fitted_and_resid_data %>% filter(model=="CES" & CSP=="Reject CSP" & (nest == "kl" | nest == "kle"))
fittedGraphNoCSP(fittedGraphData_kle)
@
%
<<Graphs_CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with $kl$ and ($le$)$k$ nesting that reject the CSP. Solid line is historical GDP.">>=
fittedGraphData_lek <- fitted_and_resid_data %>% filter(model=="CES" & CSP=="Reject CSP" & (nest == "kl" | nest == "lek"))
fittedGraphNoCSP(fittedGraphData_lek)
@
%
<<Graphs_CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with $kl$ and ($ek$)$l$ nesting that reject the CSP. Solid line is historical GDP.">>=
fittedGraphData_ekl <- fitted_and_resid_data %>% filter(model=="CES" & CSP=="Reject CSP" & (nest == "kl" | nest == "ekl"))
fittedGraphNoCSP(fittedGraphData_ekl)
@
